Thermodynamic fragments in magic state theories

### 2020-07-20 magic_paper.tex
Added analysis

Denote by $\Phi(m; n, p)$ the cumulative binomial distribution,
\begin{align}
	\Phi(m; n, p) &\coloneqq \sum\limits_{j=0}^m \binom{n}{j} p^j (1-p)^{n-j}, \\
	\Phi^c(m; n, p) &\coloneqq \sum\limits_{j=0}^m \binom{n}{j} (-p)^j (1-p)^{n-j}.
\end{align}
Note that $\Phi(m; n, p)^c$ differs only by a factor of $(-1)^j$.
Let $\Phi_+(m; n, p)$ and $\Phi_-(m; n, p)$ denote the partial sums of even and odd indices respectively,
\begin{align}	
	\Phi_+(m; n, p) &= \frac{1}{2} \left( \Phi(m; n, p) + \Phi^c(m; n, p) \right), \\
	\Phi_-(m; n, p) &= \frac{1}{2} \left( \Phi(m; n, p) - \Phi^c(m; n, p) \right).
\end{align}

\nick{$\Phi$ is bounded by the two exponentials in Wikipedia ONLY for $m\leq np$ so these bounds do not hold for $m>np$.

$\Phi^c$ is NOT bounded above by setting $m=n$ (which yields $(1 - 2p)^n$) and NOT bounded below by 0 - both due to the existence of negative terms in the sum.
It does appear however to be strictly within the open interval $(-1, 1)$.}

@@

Clearly by definition, 
\begin{align}
	\Phi_+(2m+1;n,p) &= \Phi_+(2m;n,p), \\
	\Phi_-(2m;n,p) &= \Phi_-(2m-1;n,p),
\end{align}
so we define the piece-wise functions
\begin{equation}
	L(m;n) =
	\begin{cases}
	L_+(m; n)&,\ m\text{: even}, \\
	L_-(m; n)&,\ m\text{: odd}
	\end{cases}
\end{equation}
and
\begin{equation}
	x(m;n) =
	\begin{cases}
	x_+(m; n)&,\ m\text{: even}, \\
	x_-(m; n)&,\ m\text{: odd}
	\end{cases}
\end{equation}
Finally, the $k$-th elbow is located at
\begin{equation}
	\left(x_k, \lc{\ketbra{\rm{S}}^{\otimes n}}{\frac{1}{d}\id}(k) \right) = \left( x(2k;n), L(2k;n) \right),
\end{equation}
where $2k$ is considered$\mod{(n+1)}$.

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{figs/Phi.pdf}
    \caption{Distributions $\Phi(m;n,p)$ (blue) and $\Phi^c(m;n,p)$ (orange) for $n=20$ and $p=\frac{4}{5}$.
    The green curve is the upper bound on $\Phi$, valid for all $m \leq np = 8$.
    }
    \label{fig:Phi}
\end{figure}

@@

>>>>

Denote by $\Phi(m; n, p)$ the cumulative binomial distribution,
\begin{align}
	\Phi(m; n, p) &\coloneqq \sum\limits_{j=0}^m \binom{n}{j} p^j (1-p)^{n-j}, \\
	\Phi^c(m; n, p) &\coloneqq \sum\limits_{j=0}^m \binom{n}{j} (-p)^j (1-p)^{n-j}.
\end{align}
Note that $\Phi(m; n, p)^c$ differs only by a factor of $(-1)^j$.
It is clear that for any $n,p$ and $m \leq n$
\begin{align}
	&\Phi(m; n, p) + \Phi(n-m; n, 1-p) = 1, \label{eq:reversef}\\
	&\Phi^c(m; n, p) \leq \Phi(m; n, p),
\end{align}
where equality in the second line holds iff $m=0$.

Let $\Phi_+(m; n, p)$ and $\Phi_-(m; n, p)$ denote the partial sums of even and odd indices respectively,
\begin{align}	
	\Phi_+(m; n, p) &= \frac{1}{2} \left( \Phi(m; n, p) + \Phi^c(m; n, p) \right), \\
	\Phi_-(m; n, p) &= \frac{1}{2} \left( \Phi(m; n, p) - \Phi^c(m; n, p) \right).
\end{align}

@@

Clearly by definition, 
\begin{align}
	\Phi_+(2m+1;n,p) &= \Phi_+(2m;n,p), \\
	\Phi_-(2m;n,p) &= \Phi_-(2m-1;n,p),
\end{align}
so we define the piece-wise functions
\begin{equation}
	L(m;n) =
	\begin{cases}
	L_+(m; n),&\ m\text{: even}, \\
	L_-(m; n),&\ m\text{: odd}
	\end{cases}
\end{equation}
and
\begin{equation}
	x(m;n) =
	\begin{cases}
	x_+(m; n),&\ m\text{: even}, \\
	x_-(m; n),&\ m\text{: odd}
	\end{cases}
\end{equation}
Finally, the $m$-th elbow is located at
\begin{equation}
	\left(x_m, \lc{\ketbra{\rm{S}}^{\otimes n}}{\frac{1}{d}\id}(m) \right) = \left( x(2m;n), L(2m;n) \right),
\end{equation}
where $2m$ is considered \hspace{-7pt}$\mod{(n+1)}$.

We now want to bound these points for any index $m$, so as to investigate the asymptotic behaviour of the Lorenz curve.
We define the upper bound
\begin{equation}
	U(m;n,p) \coloneqq 
	\begin{cases}
	e^{-nS\left(\frac{m}{n}||p \right)} \label{eq:ubd},&\ m \leq np, \\
	1,&\ m>np.
	\end{cases}
\end{equation}
\nick{we need to replace 1 with an exponential decay}
where 
\begin{equation}
	S(\alpha||\beta) \coloneqq \alpha \log{\frac{\alpha}{\beta}} + (1-\alpha) \log{\frac{1-\alpha}{1-\beta}},
\end{equation}
is the classical relative entropy.

There are two useful lower bounds for $\Phi(m;n,p)$, which we define as
\begin{equation}
	L(m;n,p) \coloneqq \max{\left\{ c(m;n) U(m;n,p), \\ 1 - U(n-m;n,1-p) \right\}} \label{eq:lbd},
\end{equation}
where
\begin{equation}
	c(m;n) = \left[8m\left(1-\frac{m}{n}\right)\right]^{-1/2}.
\end{equation}
Generally, the first expression within the $\max$ function is useful iff $m \leq np$, while the second expression is due to~\cref{eq:reversef} and thususeful iff $m \geq np$.
So we choose $L(m;n,p$ to be the maximum of the two at any $m$.
Figure~\cref{fig:Phi} demonstrates the upper and lower bounds for $\Phi(m;n,\frac{4}{5}$.

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{figs/Phi.pdf}
    \caption{Distributions $\Phi(m;n,p)$ (blue) and $\Phi^c(m;n,p)$ (orange) for $n=30$ and $p=\frac{4}{5}$.
    The upper and lower bounds on $\Phi$ are also presented. 
    The upper bound is given by~\cref{eq:ubd} and the lower bound by~\cref{eq:lbd}.
    }
    \label{fig:Phi}
\end{figure}

We see in~\cref{fig:phicmax} that distribution $\Phi^c(m;n,p)$ is also bounded by a constant number $\epsilon$, $0<\epsilon<1$ for all $m$.
The bound $\epsilon$ tends to decrease as $n$ increases, so that $\abs{\Phi^c(m;n,p)} \leq \frac{1}{10}$ for $n\geq 30$.

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{figs/Phic_max.pdf}
    \caption{Maximum of the absolute value of $\Phi^c(m;n,p)$ over all $m$ for given $n$.
    Here, we plot for all even $n$, $2\leq n\leq 30$.
    The maximum is decreasing in $n$ and falls below $\frac{1}{10}$ for $n=30$.
    \nick{we want to extract exponential bound from this graph instead of constant bound}.
    }
    \label{fig:phicmax}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{figs/Phi_nexp.pdf}
    \caption{Distributions $\Phi_+(m;n,p)/ U(m;n,p)$ (blue) and $U(m;n,p) \Phi_-(m;n,p)$ (orange) for $n=30$ and $p=\frac{4}{5}$.
    We multiply the bound in~\cref{eq:ubd} to extract the exponential behaviour of the two distributions.
    Here, $np = 24$, so for $m\geq 24$, $U(m;n,p) = 1$.
    }
    \label{fig:phinexp}
\end{figure}
